---
title: "p8105_hw3_qg2155"
author: "Qiu Xia (Helen) Guan"
date: "10/11/2018"
output: github_document
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r packages}
library(tidyverse)
library(patchwork)
library(p8105.datasets)
```

```{r page setup}
knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = 1.3,
  out.width = "90%"
)

theme_set(theme_bw() + theme(legend.position = "bottom"))
```

## Problem 1

### Load and clean BRFSS dataset 

```{r get brfss}
data("brfss_smart2010")
```

```{r clean brfss}
brfss_data = brfss_smart2010 %>% 
  janitor::clean_names() %>% 
  filter(topic == "Overall Health") %>% 
  mutate(response = as.factor(response))
```

### Summarizing distinct locations 

```{r states}
brfss_data %>% 
  filter(year == 2002) %>% 
  group_by(locationabbr) %>%
  summarize(n_locations = n_distinct(locationdesc))
```

In 2002, there were 3 states that were observed at 7 locations. These 3 states were CT, FL, and NC. 

### Creating spaghetti plot of number locations 

```{r spaghetti plot}
brfss_data %>% 
  group_by(locationabbr, year) %>%
  summarize(n_locations = n_distinct(locationdesc)) %>% 
  ggplot(aes(x = year, y = n_locations, color = locationabbr)) + 
    geom_point() + geom_line()
```

This spaghetti plot shows the number of locations in each state from 2002 to 2010. Florida in particular seems to have over 40 distinct loactions in two different time points. 

### Table for mean and standard deviation of "Excellent" responses in NY

```{r  excellent responses}
brfss_data %>%
  group_by(year) %>% 
  filter(year == 2002 | year == 2006 | year == 2010) %>% 
  spread(key = response, value = data_value) %>% 
  janitor::clean_names() %>% 
  filter(locationabbr == "NY") %>% 
  summarize(mean_excellent = mean(excellent, na.rm = TRUE),
            sd_excellent = sd(excellent, na.rm = TRUE)) %>% 
  knitr::kable(digits = 1)
```

Witin the state of New York, 2002 has the highest average of 24% for excellent responses, and a standard deviation of 4.5. Year 2006 has an average of 22.5% and standard deviation of 4.0. Year 2010 has an average of 22.7% and standard deviation of 3.6. 

### Plot for distribution of each response category across states

```{r five plots}
brfss_data %>% 
  group_by(year, locationabbr, response) %>%
  summarize(mean_response = mean(data_value, na.rm = TRUE)) %>% 
  ggplot(aes(x = year, y = mean_response)) +
    geom_violin() +
    facet_grid(. ~ response) 
```

Across all states and years, the response category of "Good" seems to have the highest average, followed by "Excellent", "Fair", and "Poor". Excellent seems to have the biggest range out of them all. 

## Problem 2

### Load and clean dataset

```{r get instacart}
data("instacart")
```

```{r clean instacart}
instacart_data = instacart %>% 
  janitor::clean_names() 
```

### Aisles and most items ordered 

```{r most items}
instacart_data %>% 
  group_by(aisle_id) %>% 
  summarize(n_most_ordered = n()) %>% 
  arrange(desc(n_most_ordered))
```

There are `r instacart_data %>% summarize(n_aisles = n_distinct(aisle_id))` aisles. Aisle 83 has the most item ordered from because 150609 items were ordered from that aisle. This is followed by aisle 24 with 150473 items, and aisle 123 with 78493 items. 

### Plot for number of items ordered in each aisle

```{r aisle plot}
instacart_data %>% 
  group_by(aisle_id) %>% 
  summarize(n_most_ordered = n()) %>% 
  ggplot(aes(x = n_most_ordered)) +
    geom_histogram() +
    labs(
      title = "Items ordered from aisle",
      x = "Number of items ordered from each aisle",
      y = "Count of items"
    )
```

### Table for most popular item in each of the three aisles

```{r popular item}
instacart_data %>% 
  group_by(aisle, product_name) %>% 
  filter(aisle %in% c("baking ingredients", "dog food care", "packaged vegetables fruits")) %>% 
  summarize(n_ordered = n()) %>% 
  arrange(desc(n_ordered)) %>%
  top_n(1) %>% 
  knitr::kable()
```

The most popular product in the "packaged vegetables fruits" aisle is Organic Baby Spinach. It was ordered 9784 times. The most popular product in the "baking ingredients" aisle is Light Brown Sugar. It was ordered 499 times. The most popular product in the "dog food care" aisle is Snack Sticks Chicken & Rice Recipe Dog Treats. It was ordered 30 times. 

### Table for Pink Lady Apples and Coffee Ice Cream Order Times

```{r}
instacart_data %>% 
  select(order_id, product_id, order_dow, order_hour_of_day, product_name, aisle_id) %>% 
  filter(product_name %in% c("Pink Lady Apples", "Coffee Ice Cream")) %>% 
  group_by(product_name, order_dow) %>%
  summarize(mean_order_hour = mean(order_hour_of_day)) %>% 
  spread(key = order_dow, value = mean_order_hour) %>% 
  knitr::kable(digits = 2)
```

The average order time for Coffee Ice Cream ranges from 12:26 to 15:38 throughout the week. The day with the earliest average order time is Friday and the day with the latest average order time is Tuesday. Pink Lady Apples have an overall earlier average order time than Coffee Ice Cream. The range for Pink Lady Apples average order times are 11:36 to 14:25. The day with the earliest average order time is Monday and the day with the latest average order time is Wednesday. 


